High-Dimensional Regularized Discriminant Analysis

### 2016-10-23 reports/paper-hdrda.Rnw
Revamped abstract #28

\citeauthor{Friedman:1989tm} proposed the popular regularized discriminant
analysis (RDA) classifier that utilizes a biased covariance-matrix estimator
that partially pools the sample covariance matrices from linear and quadratic
discriminant analysis and shrinks the resulting estimator towards a scaled
identity matrix. The RDA classifier's two tuning parameters are typically
estimated via a computationally burdensome cross-validation procedure that uses
a grid search. We formulate a new RDA-based classifier for the small-sample,
high-dimensional setting and then show that the classification decision rule is
equivalent to a classifier in a subspace having a much lower dimension. As a
result, the utilization of the dimension-reduction step yields a substantial
reduction in computation during model selection. Also, our parameterization
offers interpretability that was previously lacking with the RDA classifier. We
demonstrate that our proposed classifier is often superior to several recently
proposed sparse and regularized classifiers in terms of classification accuracy
with three artificial and six real high-dimensional data sets. Finally, we
provide an implementation of our proposed classifier in the {\tt sparsediscrim}
R package, which is available on CRAN.

>>>>

We develop a new and computationally efficient implementation of
\citeauthor{Friedman:1989tm}'s popular regularized discriminant analysis (RDA)
classifier in the small-sample, high-dimensional setting. Our proposed
high-dimensional RDA (HDRDA) classifier modifies \citeauthor{Friedman:1989tm}'s
original formulation to achieve interpretability previously lacking. We are able
to interpret the biased covariance matrix estimator employed in the HDRDA
classifier as one that is estimated from all training observations regardless of
class, solidifying the ``borrowing'' operation \citeauthor{Friedman:1989tm}'s
RDA is claimed to have. Moreover, our formulation also differs from
\citeauthor{Friedman:1989tm}'s in that we show our classifier is equivalent to
one in a reduced feature space with dimension approximately equal to the
training sample size. As a result, the matrix operations used in the HDRDA
classifier are computationally linear in the number of features, making HDRDA
well-suited for high-dimensional classification in practice. Timing comparisons
between our HDRDA implementation in the {\tt sparsediscrim} R package and the
standard RDA formulation in the {\tt klaR} R package demonstrate that as the
number of features increases, the computational runtime of HDRDA is drastically
smaller than RDA. We also demonstrate that HDRDA is often superior to several
sparse and regularized classifiers in terms of classification accuracy with
three artificial and six real high-dimensional data sets.
